Embedding Intuitionistic into Classical Logic

### 2023-03-07 LPAR/LPAR-EIICL.tex
Syntax redefined, Proof 7.2 to appendix.

For propositional logic we assume an ambient set of propositional variables. A propositional formula is defined inductively as follows:
\begin{itemize}
	\item The propositional constant $\bot$ is a formula.
	\item Each propositional variable is a formula.
	\item Given formulas $\varphi$ and $\psi$ and a connector $\circ\in\{\wedge, \vee,\to\}$ there is a formula $\varphi\circ\psi$.
\end{itemize}

For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and sets of free and bound variables. Generally free variables will be denoted will letters at the beginning of the alphabet while bound variables will be denoted with letters at the end of the alphabet. $\nu$ will usually be a placeholder that can be instantiated with free or bound variables. We can define terms $t$ and substitutions $t[\nu/b]$ by simultaneous induction as follows:
\vspace*{-.3cm}
\begin{itemize}
	\item Each free variable $a$ is a term and $a[\nu/b] := \begin{cases}
		\nu & \text{if } a = b \\
		a & \text{otherwise}
	\end{cases}$.
	\item For each $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$ we have that $f(t_1, \ldots, t_n)$ is a term and $f(t_1, \ldots, t_n)[\nu/b] := f(t_1[\nu/b], \ldots, t_n[\nu/b])$.
\end{itemize}
We can then define formulas $\varphi$ and substitutions $\varphi[\nu/b]$ by simultaneous induction as follows:
\begin{itemize}
	\item The constant $\bot$ is a formula and $\bot[\nu/b] := \bot$.
	\item For each $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$ we have that $P(t_1, \ldots, t_n)$ is a formula and $P(t_1,\ldots, t_n)[\nu/b] := P(t_1[\nu/b], \ldots, t_n[\nu/b])$.
	\item Given formulas $\varphi$ and $\psi$ and a connector $\circ\in \{\wedge,\vee,\to\}$ there is a formula $\varphi\circ\psi$ and $(\varphi\circ\psi)[\nu/b] := \varphi[\nu/b]\circ\psi[\nu/b]$.
	\item Given a formula $\varphi$, a free variable $a$, a bound variable $x$ not occurring in $\varphi$ and a quantifier $Q\in\{\exists,\forall\}$ there is a formula $Qx(\varphi[x/a])$ and $Qx(\varphi[x/a])[\nu/b] := Qx(\varphi[x/a][\nu/b])$.
\end{itemize}

@@

\begin{definition}
	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary predicate symbol $R\in\Sigma$ a $n$-ary predicate $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
	m, &\text{if $b=a$,}\\
	v(b), &\text{otherwise.}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$ for each $n$-ary function symbol $f\in\Sigma$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$
		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.
	\end{itemize}
	 Here $a$ is a free variable that does not occur in $\varphi$. A $\Sigma$-structure $\mathcal M$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$, we write $\mathcal M\models\varphi$. A formula is valid if every $\Sigma$-structure $\mathcal M$ satisfies it. We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
\end{definition}

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulas $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
	\begin{itemize}
		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/a]\models\varphi[a/x]$.
	\end{itemize}
	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$. $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
	$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Kripke structure with that frame. We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to propositional tautologies there are now quantified formulas which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.

@@

@@

>>>>

For propositional logic we assume an ambient set of propositional variables. As usual a formula is either $\bot$, a propositional variable or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$.

For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and a set of variables. As usual we can inductively define terms and formuas. A term is either a variable or, given an $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$, $f(t_1, \ldots, t_n)$. A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$. Free and bound variables are defined as usual and we assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables. Generally we will denote free variables with letters at the beginning of the alphabet and bound variables with letters at the end of the alphabet.

@@

\begin{definition}
	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary predicate symbol $R\in\Sigma$ a $n$-ary predicate $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
	m, &\text{if $b=a$,}\\
	v(b), &\text{otherwise.}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$ for each $n$-ary function symbol $f\in\Sigma$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$
		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.
	\end{itemize}
	 Here $a$ is a free variable that does not occur in $\varphi$. $\varphi$ is satisfiable if $\mathcal M, v\models\varphi$ for some $\mathcal M, v$. It is valid if $\mathcal M, v\models\varphi$ for every $\mathcal M, v$. We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
\end{definition}

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulas $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
	\begin{itemize}
		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/a]\models\varphi[a/x]$.
	\end{itemize}
	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$. $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
	We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to propositional tautologies there are now quantified formulas which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.

@@

A proof can be found in the appendix~\ref*{proof:fo-simplification}.

@@

\begin{lemma}\label{proof:fo-simplification}
	$\varphi$ is intuitionistically valid if and only if $\varphi^C$ is classically valid.
\end{lemma}
